|

### 2022-03-28 01-Chp-Background.Rmd
Range of corrections

[Alzheimer's disease is the most common cause of dementia, accounting for approximately 60-80% of cases.]{.correction} [@alzheimersassociation2020] Characterised by substantial cognitive impairment and difficulty with high level executive function, it is an insidious disease, with initial onset thought to occur up to 15 years prior to symptomatic presentation.[@robinson2015] Much remains unknown about Alzheimer's pathogenesis, despite research implicating the "amyloid hypothesis",[@robinson2015] as a potential mechanism of disease. Under this hypothesis, the build-up of amlyoid plaques (composed mainly of amlyoid-$\beta$ peptide) and neurofibrillary tangles (composed mainly of tau protein) triggers a range of physiological changes, including inflammation and cell death, that result in cognitive impairment.[@robinson2015]

Vascular dementia (VaD) is the second largest underlying pathology of dementia, accounting for ~10-25% of cases. Vascular dementia is caused by a range of cerebrovascular disorders, and as a result, presentation  of symptoms can vary widely.[@iadecola2013] Similarly, due to the varied underlying  pathophysiology, vascular dementia can onset either quite rapidly following a cerebrovascular event such as a stroke or over a long time-frame due to a series of small infarcts.[@venkat2015] Vascular dementia regularly co-occurs in patients with Alzheimer's disease.[@iadecola2013] This presentation is described as "mixed" dementia,[@custodio2017] and occurs in approximately 25% of cases.[@burns2009]

@@

Both trials were also limited by the relatively short follow-up period examined. This limited follow-up was to be expected because the primary aim of both trials was to assess the effect of statins on short-term coronary-related outcomes.[@heartprotectionstudycollaborativegroup2002; @trompet2010] The PROSPER trial had a mean follow-up of 3.2 years, while the MRC/BHF Protection Study estimated risk at 5 years of follow-up. Given the long lag time between non-symptomatic onset of dementia and clinical presentation, it is likely that these durations are insufficient to fully capture the onset of dementia. Finally, as they included only patients at high vascular risk, their generalisability to other settings is limited.[@mcguinness2016]

@@

```{block type = "correction"}
As part of this research, I attempted to obtain the raw data from relevant primary studies identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading). Data obtained were combined with that available from the DPUK portal as part of an IPD meta-analysis in Chapter \@ref(ipd-heading). This enabled the assessment of the effect of lipids on dementia, and whether this relationship is modified by key variables such as participants' age or sex.
```

>>>>

Alzheimer's disease is the most common cause of dementia, accounting for approximately 60-80% of [cases.]{.correction}[@alzheimersassociation2020] Characterised by substantial cognitive impairment and difficulty with high level executive function, it is an insidious disease, with initial onset thought to occur up to 15 years prior to symptomatic presentation.[@robinson2015] Much remains unknown about Alzheimer's pathogenesis, despite research implicating the "amyloid hypothesis",[@robinson2015] as a potential mechanism of disease. Under this hypothesis, the build-up of amlyoid plaques (composed mainly of amlyoid-$\beta$ peptide) and neurofibrillary tangles (composed mainly of tau protein) triggers a range of physiological changes, including inflammation and cell death, that result in cognitive impairment.[@robinson2015]

Vascular dementia (VaD) is the second largest underlying pathology of dementia, accounting for ~10-25% of [cases.]{.correction}[@alzheimersassociation2020] Vascular dementia is caused by a range of cerebrovascular disorders, and as a result, presentation  of symptoms can vary widely.[@iadecola2013] Similarly, due to the varied underlying  pathophysiology, vascular dementia can onset either quite rapidly following a cerebrovascular event such as a stroke or over a long time-frame due to a series of small infarcts.[@venkat2015] Vascular dementia regularly co-occurs in patients with Alzheimer's disease.[@iadecola2013] This presentation is described as "mixed" dementia,[@custodio2017] and occurs in approximately 25% of cases.[@burns2009]

@@

Both trials were also limited by the relatively short follow-up period examined. This limited follow-up was to be expected because the primary aim of both trials was to assess the effect of statins on short-term coronary-related outcomes.[@heartprotectionstudycollaborativegroup2002; @trompet2010] The PROSPER trial had a mean follow-up of 3.2 years, while the MRC/BHF Protection Study estimated risk at 5 years of follow-up. Given the long lag time between non-symptomatic onset of dementia and clinical presentation, it is likely that these durations are insufficient to fully capture the onset of dementia. [Further]{.correction}, as they included only patients at high vascular risk, their generalisability to other settings is limited.[@mcguinness2016]

[Finally, the ability of studies of statin use to assess a causal link between lipid levels and dementia is limited by the pleiotropic effect of the intervention, as statins may effect dementia risk via some non-lipid-lowering pathway. For example, statins influence the levels of C-reactive protein,]{.correction}[@albert2001] [an independent risk factor for dementia.]{.correction}[@lewis2021; @schmidt2002]

@@

[As part of this research, I attempted]{.correction} to obtain the raw data from relevant primary studies identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading). Data obtained [were]{.correction} combined with that available from the DPUK portal as part of an IPD meta-analysis in Chapter \@ref(ipd-heading). This [enabled]{.correction} the assessment of the effect of lipids on dementia, and whether this relationship is modified by key variables such as participants' age or sex.

### 2022-03-28 03-Chp-Systematic-Review-Methods.Rmd
Range of corrections

I sought to include studies that examined blood lipid levels as a risk factor for dementia outcomes, defined either as a binary hypercholesterolemia variable or by category/1-standard-deviation increase in a specific lipid fraction. The fractions considered by this review were namely total cholesterol (TC), high-density lipoprotein cholesterol (HDL), low-density lipoprotein cholesterol (LDL) and triglycerides (TG). I also aimed to include studies examining the effect of lipid-regulating agents (LRA) as a source of indirect evidence. LRA are treatments used to modify blood lipids levels, and common examples include statins, fibrates, and ezetimibe.

@@

Studies of any duration were included, and no limits were placed on the sample size of included studies. Finally, conference abstracts with no corresponding full-text publication were eligible, and where required, I contacted authors to obtain relevant additional information not available from the abstract. No limitations were imposed on publication status, date, venue or language.

@@

An initial qualitative synthesis of evidence was performed, summarising the data extracted from studies stratified by study design. Where individual studies were deemed comparable, they were incorporated into a quantitative analysis or "meta-analysis". Results were not combined across different study designs (i.e., RCTs were not combined in a meta-analysis with results from observational studies). The summary effect estimates produced across individual study designs are discussed but are compared and contrasted more fully as part of the triangulation exercise presented in Chapter \@ref(tri-heading). Similarly, analyses are presented separately for each dementia outcome of interest (all-cause dementia, Alzheimer's disease, vascular dementia).

>>>>

I sought to include studies that examined blood lipid levels as a risk factor for dementia outcomes, defined either as a binary hypercholesterolemia variable or by category/1-standard-deviation increase in a specific lipid fraction. The fractions considered by this review were namely total cholesterol (TC), high-density lipoprotein cholesterol [(HDL-c)]{.correction}, low-density lipoprotein cholesterol [(LDL-c)]{.correction} and triglycerides (TG). I also aimed to include studies examining the effect of lipid-regulating agents (LRA) as a source of indirect evidence. LRA are treatments used to modify blood lipids levels, and common examples include statins, fibrates, and ezetimibe.

@@

[Longitudinal]{.correction} studies [(randomised trials and observational studies)]{.correction} of any duration were included, and no limits were placed on the sample size of included studies. Finally, conference abstracts with no corresponding full-text publication were eligible, and where required, I contacted authors to obtain relevant additional information not available from the abstract. No limitations were imposed on publication status, date, venue or language.

@@

An initial qualitative synthesis of evidence was performed, summarising the data extracted from studies stratified by study design. Where individual studies were deemed comparable, they were incorporated into a quantitative analysis or "meta-analysis". Results were not combined across different study designs (i.e., RCTs were not combined in a meta-analysis with results from observational studies). The summary effect estimates produced across individual study designs are discussed but are compared and contrasted more fully as part of the triangulation exercise presented in Chapter \@ref(tri-heading). Similarly, analyses are presented separately for each dementia outcome of interest (all-cause dementia, Alzheimer's disease, vascular dementia). [A final division of analyses was made based on the reported exposure measurement (e.g. binary measurement of hypercholesterolemia, per standard deviation change in lipid fraction, etc.).]{.correction}

### 2022-03-28 05-Chp-CPRD-Analysis.Rmd
Range of corrections

This finding disagrees with the results of the original analysis, which found evidence for a protective effect of statin use on all-cause dementia (`r estimate(0.81, 0.69,0.96,type = "HR")`) and non-AD  dementia (`r estimate(0.82,0.69, 0.97,type = "HR")`), but little evidence of an effect on AD (`r estimate(0.81,0.49, 1.35, type="HR")`).

>>>>

This finding disagrees with the results of the original analysis [reported by Smeet _et al_,]{.correction} [@smeeth2009] which found evidence for a protective effect of statin use on all-cause dementia (`r estimate(0.81, 0.69,0.96,type = "HR")`) and non-AD dementia (`r estimate(0.82,0.69, 0.97,type = "HR")`), but little evidence of an effect on AD (`r estimate(0.81,0.49, 1.35, type="HR")`).
